---
title: "2024-04-30-Considering the Prospect of Rising Interest Rates"
author: "Edith"
date: "2024-04-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r , include=FALSE}

library(dplyr)
library(xts)
library(ggplot2)
library(lubridate);
library(fredr)

fredr_set_key("76e5a66429cbd1b05fbda4b2eeaf3544")
```


#### Will the Federal Reserve raise interest rates?

My view leans towards rates either holding steady or increasing. The primary rationale for this expectation is that there are no immediate signs suggesting the economy is on the verge of a recession. 
The labor market remains robust, the Conference Board's Consumer Confidence Index continues to show strength, and inflation has stabilized around 3%. These indicators suggest that the U.S. economy is not currently in a downturn, nor does it appear to be heading towards one.

However, the risks associated with maintaining high interest rates cannot be overlooked, as they might lead to an economic slowdown over time. Despite these concerns, if this were the sole reason for the Fed to consider lowering rates, it is unlikely they would act on that basis alone. It is more probable that they will adopt a 'wait and see' approach, evaluating whether the economy can sustain its growth under the current interest rate regime.


```{r pressure}
library(dplyr)
library(xts)
library(ggplot2)
library(lubridate);
library(fredr)
# Set your FRED API key
# fredr_set_key("your key")

# Fetch the data
inf_us <- fredr(
  series_id = "CPIAUCNS",
  frequency = "m",
  units = "pc1" # "pc1" - Percent change from 1 year ago
)
unemp_us <- fredr(
  series_id = "UNRATE",
  frequency = "m" # m: monthly, q: quarterly
)

# Plot the data
ggplot(inf_us[inf_us$date >= as.Date("2000-01-01"), ]) +
  geom_line(mapping = aes(x = date, y = value), color = "deepskyblue3", linewidth = 0.6) +
  ggtitle("US CPI Inflation") +
  xlab("Year") +
  ylab("%")


ggplot(unemp_us[unemp_us$date >= as.Date("2000-01-01"), ]) +
  geom_line(mapping = aes(x = date, y = value), color = "red") +
  ggtitle("US Unemployment rate") +
  xlab("Year") +
  ylab("%")
```


